---
title: "Austin Growth Model"
author: "Adil Belgaumi and Mimi Tran"
date: "2024-04-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, message=FALSE, warning=FALSE, results = "hide"}
if(!require(pacman)){install.packages("pacman"); library(pacman)}
p_load(tidyverse, sf, raster, knitr, kableExtra, tidycensus, tigris, FNN, caret,
       yardstick, pscl, plotROC, ggrepel, pROC, grid, gridExtra, viridis, igraph,
       tmap,rasterVis)

plotTheme <- theme(
  plot.title =element_text(size=12),
  plot.subtitle = element_text(size=8),
  plot.caption = element_text(size = 6),
  axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
  axis.text.y = element_text(size = 10),
  axis.title.y = element_text(size = 10),
  # Set the entire chart region to blank
  panel.background=element_blank(),
  plot.background=element_blank(),
  #panel.border=element_rect(colour="#F0F0F0"),
  # Format the grid
  panel.grid.major=element_line(colour="#D0D0D0",size=.75),
  axis.ticks=element_blank())

mapTheme <- theme(plot.title =element_text(size=12),
                  plot.subtitle = element_text(size=8),
                  plot.caption = element_text(size = 6),
                  axis.line=element_blank(),
                  axis.text.x=element_blank(),
                  axis.text.y=element_blank(),
                  axis.ticks=element_blank(),
                  axis.title.x=element_blank(),
                  axis.title.y=element_blank(),
                  panel.background=element_blank(),
                  panel.border=element_blank(),
                  panel.grid.major=element_line(colour = 'transparent'),
                  panel.grid.minor=element_blank(),
                  legend.direction = "vertical", 
                  legend.position = "right",
                  plot.margin = margin(1, 1, 1, 1, 'cm'),
                  legend.key.height = unit(1, "cm"), legend.key.width = unit(0.2, "cm"))

palette2 <- c("#41b6c4","#253494")
palette4 <- c("#a1dab4","#41b6c4","#2c7fb8","#253494")
palette5 <- c("#ffffcc","#a1dab4","#41b6c4","#2c7fb8","#253494")
palette10 <- c("#f7fcf0","#e0f3db","#ccebc5","#a8ddb5","#7bccc4",
               "#4eb3d3","#2b8cbe","#0868ac","#084081","#f7fcf0")
```

```{r Functions}
quintileBreaks <- function(df,variable) {
    as.character(quantile(df[[variable]],
                          c(.01,.2,.4,.6,.8),na.rm=T))
}

#This function can be used to convert a polygon sf to centroids xy coords.
xyC <- function(aPolygonSF) {
  as.data.frame(
    cbind(x=st_coordinates(st_centroid(aPolygonSF))[,1],
          y=st_coordinates(st_centroid(aPolygonSF))[,2]))
} 

#this function convert a raster to a data frame so it can be plotted in ggplot
rast <- function(inRaster) {
  data.frame(
    xyFromCell(inRaster, 1:ncell(inRaster)), 
    value = getValues(inRaster)) }
```


```{r}
acs_variables <- load_variables(2021, "acs5")

MSAShape <- get_acs(
  geography = "county", variables = "B01001_001", state = 48,  geometry = TRUE, 
                county=c("Hays", "Travis", "Williamson","Bastrop","Caldwell"), year = 2021) %>% 
  st_transform(crs = 6578) %>%
  st_union()

studyAreaCounties <- 
  counties("Texas") %>%
  st_transform(st_crs(MSAShape)) %>%
  dplyr::select(NAME) %>%
  .[st_buffer(MSAShape,-1000), , op=st_intersects]

ggplot() +
  geom_sf(data=studyAreaCounties) +
  labs(title = "Study Area Counties") +
  mapTheme

studyAreaCounties <- st_as_sf(studyAreaCounties)

pop_2021 <- get_acs(
  geography = "tract", variables = "B01001_001", state = 48,  geometry = TRUE, 
                county=c("Hays", "Travis", "Williamson","Bastrop","Caldwell"), year = 2021) %>% 
  st_transform(crs = 6578) %>%
  rename(pop_2021 = estimate) %>% 
  st_buffer(-1)

pop_2011 <- get_acs(
  geography = "tract", variables = "B01001_001", state = 48,  geometry = TRUE, 
                county=c("Hays", "Travis", "Williamson","Bastrop","Caldwell"), year = 2011) %>% 
  st_transform(crs = 6578) %>%
  rename(pop_2011 = estimate) %>% 
  st_buffer(-1)

tm_shape(pop_2011) +
  tm_fill(col = "pop_2011")
# st_write(pop_2021,dsn = "Data/pop_2021.shp", append = FALSE)
```

Creating the Fishnet
```{r}
Austin_fishnet <- 
  st_make_grid(MSAShape, 4000) %>%
  st_sf() %>% 
  .[pop_2021, , op=st_intersects]
```


```{r Census Data}
pop_fishnet <-
  Austin_fishnet %>%
  rownames_to_column("fishnetID") %>% 
  mutate(fishnetID = as.numeric(fishnetID)) %>%
  dplyr::select(fishnetID)

fishnetPop11 <-
  st_interpolate_aw(pop_2011["pop_2011"], pop_fishnet, extensive=TRUE) %>%
  as.data.frame(.) %>%
  rownames_to_column(var = "fishnetID") %>%
  left_join(pop_fishnet %>%
              mutate(fishnetID = as.character(fishnetID)),
            ., by=c("fishnetID"='fishnetID')) %>% 
  mutate(pop_2011 = replace_na(pop_2011,0)) %>%
  dplyr::select(pop_2011)

fishnetPop21 <-
  st_interpolate_aw(pop_2021["pop_2021"], pop_fishnet, extensive=TRUE) %>%
  as.data.frame(.) %>%
  rownames_to_column(var = "fishnetID") %>%
  left_join(pop_fishnet %>%
              mutate(fishnetID = as.character(fishnetID)),
            ., by=c("fishnetID"='fishnetID')) %>% 
  mutate(pop_2010 = replace_na(pop_2021,0)) %>%
  dplyr::select(pop_2021)

pop_fishnet <- 
  cbind(fishnetPop11,fishnetPop21) %>%
  dplyr::select(pop_2011,pop_2021) %>%
  mutate(pop_Change = pop_2021 - pop_2011)

tmap_arrange(
   tm_shape(pop_fishnet) +
     tm_fill(col = "pop_2011", palette = palette5, n = 5, style = "quantile") +
     tm_layout(legend.position = c("right", "bottom")),
   tm_shape(pop_fishnet) +
     tm_fill(col = "pop_2021", palette = palette5, n = 5, style = "quantile") +
     tm_layout(legend.position = c("right", "bottom")),
   tm_shape(pop_fishnet) +
     tm_fill(col = "pop_Change", palette = palette5, n = 5, style = "quantile") +
     tm_layout(legend.position = c("right", "bottom")),
   ncol = 2)
```


```{r}
if(Sys.getenv("USERNAME") == "adgob") {
  lc_austin_11 <- raster("Data/lc_final/lc_austin_11.tif")
  lc_austin_21 <- raster("Data/lc_final/lc_austin_21.tif")
}  else {
    #put your directory paths here !
   lc_austin_11 <- raster("~/Desktop/modeling/HW5/austin 2021/lc_austin_21.tif")
   lc_austin_21 <- raster("~/Desktop/modeling/HW5/austin 2011/lc_austin_11.tif")
}
```

```{r Reclass Matrices}
devM <- 
  matrix(c(
    0,12,0,
    12,24,1,
    24,Inf,0),
  ncol=3, byrow=T)

forestM <- 
  matrix(c(
    0,40,0,
    40,45,1,
    45,Inf,0),
    ncol=3, byrow=T
  )

farmM <- 
  matrix(c(
    0,80,0,
    80,90,1,
    90,Inf,0),
    ncol=3, byrow=T
  )

wetlandM <- 
  matrix(c(
    0,89,0,
    89,100,1,
    100,Inf,0),
    ncol=3, byrow=T
  )


oUndevM <- 
  matrix(c(
    0,30,0,
    30,35,1,
    35,50,0,
    50,55,1,
    55,70,0,
    70,75,1,
    75,Inf,0),
    ncol=3, byrow=T
  )

waterM <- 
  matrix(c(
    0,15,1,
    15,Inf,0),
    ncol=3, byrow=T
  )

ClassMatrices <- list(devM,forestM,farmM,wetlandM,oUndevM,waterM)
ClassNames <- list("dev","forest","farm","wetland","oUndev","water")

```


```{r}
Raster_list2011 <- c()
Raster_names <- c()
for (x in 1:length(ClassMatrices)){ 
  # Reclassify land cover to the desired land cover type
  data <- reclassify(lc_austin_11, ClassMatrices[[x]])
  # Aggregate to get fraction of cell with that land type
  data <- aggregate(data, fact = 40, FUN = mean)
  # Name the data in the raster x_2011
  names(data) <- ClassNames[[x]]
  # Adds the raster to the raster list with the name x_2011_4000
  Raster_list2011 <- c(Raster_list2011,
                assign(ClassNames[[x]],data))
  # Adds the raster layer names to a list for running future code
  Raster_names <-  c(Raster_names,
                     ClassNames[[x]])
}

dev_2021 <- reclassify(lc_austin_21, devM)
dev_2021_4000 <- aggregate(dev_2021, fact = 40, FUN = mean)
names(dev_2021_4000) <- "dev_2021"
Raster_list2011 <- c(Raster_list2011, 
                     dev_2021_4000)
Raster_names <-  c(Raster_names,
                     "dev_2021")

development_change <- dev_2021_4000 - Raster_list2011[[1]] 
names(development_change) <- "dev_change"
Raster_list2011 <- c(Raster_list2011, 
                     development_change)
Raster_names <-  c(Raster_names,
                     "dev_change")
```


```{r}

# lc_au_11_2000 <- aggregate(lc_austin_11, fact = 40, FUN= max)
# lc_au_21_2000 <- aggregate(lc_austin_21, fact = 40, FUN= max)
# 
# ggplot() +
#   geom_raster(data=rast(Raster_list2011[[1]]) %>% na.omit %>% filter(value > 0),
#               aes(x,y,fill=value)) +
#   labs(title = "Land Cover, 2011") +
#   mapTheme +
#   theme(legend.direction="vertical")
```

```{r Aggregate Raster to Fishnet Function, warning = FALSE, message = FALSE}
aggregateRaster <- function(inputRasterList, theFishnet) {
  #create an empty fishnet with the same dimensions as the input fishnet
  theseFishnets <- theFishnet %>% dplyr::select()
  #for each raster in the raster list
  for (i in inputRasterList) {
  #create a variable name corresponding to the ith raster
  varName <- names(i)
  #convert raster to points as an sf
    thesePoints <-
      rasterToPoints(i) %>%
      as.data.frame() %>%
      st_as_sf(coords = c("x", "y"), crs = st_crs(theFishnet)) %>%
      filter(.[[1]] > 0)
  #aggregate to the fishnet
    thisFishnet <-
      aggregate(thesePoints, theFishnet, mean) %>%
      mutate(!!varName := ifelse(is.na(.[[1]]),0,.[[1]]))
 #add to the larger fishnet
    theseFishnets <- cbind(theseFishnets,thisFishnet)
  }
  #output all aggregates as one large fishnet
   return(theseFishnets)
}
```

```{r}
aggregatedRasters <-
  aggregateRaster(Raster_list2011, Austin_fishnet) %>%
  dplyr::select(all_of(Raster_names))
```

```{r}
aggregatedRasters <- aggregatedRasters %>% 
  mutate(developed2011 = ifelse(dev > 0.25,1,0), # made a difference between what was considered development in 2011 and 2021 
         developed2021 = ifelse(dev_2021 > 0.2,1,0),
         development_change2 = developed2021 - developed2011
         ,development_change2 = ifelse(development_change2 < 1,
                                      0,
                                      1),
         maj_water = water > 0.5 # was 0.5 
         )

aggregatedRasters %>% 
  group_by(maj_water,developed2011,developed2021) %>% 
  tally()


tm_shape(aggregatedRasters) +
  tm_fill(col = "developed2011", palette = palette2)
tm_shape(aggregatedRasters) +
  tm_fill(col = "developed2021", palette = palette2)
tm_shape(aggregatedRasters) +
  tm_fill(col = "development_change2", palette = palette5, n = 5, style = "quantile")

```


```{r}
tmap_arrange(
  tm_shape(aggregatedRasters) +
  tm_fill(col = "dev", palette = palette5, n = 5, style = "quantile") +
    tm_layout(legend.position = c("right", "bottom")),
  tm_shape(aggregatedRasters) +
    tm_fill(col = "forest", palette = palette5, n = 5, style = "quantile") +
    tm_layout(legend.position = c("right", "bottom")),
  tm_shape(aggregatedRasters) +
    tm_fill(col = "water", palette = palette5, n = 5, style = "quantile") +
    tm_layout(legend.position = c("right", "bottom")),
  tm_shape(aggregatedRasters) +
    tm_fill(col = "oUndev", palette = palette5, n = 5, style = "quantile") +
    tm_layout(legend.position = c("right", "bottom")), 
  tm_shape(aggregatedRasters) +
    tm_fill(col = "farm", palette = palette5, n = 5, style = "quantile") +
    tm_layout(legend.position = c("right", "bottom")), 
  tm_shape(aggregatedRasters) +
    tm_fill(col = "wetland", palette = palette5, n = 5, style = "quantile") +
    tm_layout(legend.position = c("right", "bottom")),
  tm_shape(aggregatedRasters) +
    tm_fill(col = "dev_change", palette = palette5, n = 5, style = "quantile") +
    tm_layout(legend.position = c("right", "bottom")), 
  tm_shape(aggregatedRasters) +
    tm_fill(col = "maj_water", palette = palette5, n = 5, style = "quantile") +
    tm_layout(legend.position = c("right", "bottom")), 
  ncol = 2)
```


```{r Highway Data load}
if(Sys.getenv("USERNAME") == "adgob") {
  MSAHighways <- st_read("Data/TxDOT_Roadways.shp")
}  else {
    #put your directory paths here !
   MSAHighways <- st_read("~/Desktop/modeling/HW5/TxDOT_Roadways_1642518016996364346/TxDOT_Roadways.shp")
}


MSAHighways <- MSAHighways  %>% 
  filter(RTE_PRFX %in% c("IH", 
                         "SH", 
                         "SA",
                         "SL",
                         "UA",
                         "US",
                         "TL"
                         )) %>% 
  st_transform(st_crs(pop_2021)) %>%  
  st_zm() %>% 
  .[pop_2021, , op=st_intersects]
```


```{r Highway Data to Fishnet}
emptyRaster <- dev_2021_4000
emptyRaster[] <- NA

highway_raster <- 
  as(MSAHighways,'Spatial') %>%
  rasterize(.,emptyRaster)

highway_raster_distance <- distance(highway_raster)
names(highway_raster_distance) <- "distance_highways"

highwayPoints <-
  rasterToPoints(highway_raster_distance) %>%
  as.data.frame() %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(Austin_fishnet))

highwayPoints_fishnet <- 
  aggregate(highwayPoints, Austin_fishnet, mean) %>%
  mutate(distance_highways = ifelse(is.na(distance_highways),0,distance_highways))

# ggplot() +
#   geom_point(data=highwayPoints_fishnet, aes(x=xyC(highwayPoints_fishnet)[,1], 
#                                              y=xyC(highwayPoints_fishnet)[,2], 
#                  colour=factor(ntile(distance_highways,5))),size=1.5) +
#   scale_colour_manual(values = palette5,
#                       labels=substr(quintileBreaks(highwayPoints_fishnet,"distance_highways"),1,8),
#                       name="Quintile\nBreaks") +
#   geom_sf(data=MSAHighways, colour = "red") +
#   labs(title = "Distance to Highways",
#        subtitle = "As fishnet centroids; Highways visualized in red") +
#   mapTheme

tm_shape(highwayPoints_fishnet) +
  tm_fill(col = "distance_highways", palette = palette5, n = 5, style = "quantile")
```


```{r Development Lag Calc}
nn_function <- function(measureFrom,measureTo,k) {
  #convert the sf layers to matrices
  measureFrom_Matrix <-
    as.matrix(measureFrom)
  measureTo_Matrix <-
    as.matrix(measureTo)
  nn <-   
    get.knnx(measureTo, measureFrom, k)$nn.dist
    output <-
    as.data.frame(nn) %>%
    rownames_to_column(var = "thisPoint") %>%
    gather(points, point_distance, V1:ncol(.)) %>%
    arrange(as.numeric(thisPoint)) %>%
    group_by(thisPoint) %>%
    summarize(pointDistance = mean(point_distance)) %>%
    arrange(as.numeric(thisPoint)) %>% 
    dplyr::select(-thisPoint) %>%
    pull()
  
  return(output)  
}

Austin_fishnet$lagDevelopment <-
    nn_function(xyC(Austin_fishnet),
                xyC(filter(aggregatedRasters,dev>0.2)), 
                2) # was 0.3

Austin_fishnet$lagMajDevelopment <-
    nn_function(xyC(Austin_fishnet),
                xyC(filter(aggregatedRasters,dev>0.8)),
                2)
```

## Final Dataset

```{r}

dat <- 
  cbind(
    Austin_fishnet$lagDevelopment,
    aggregatedRasters, 
    st_drop_geometry(highwayPoints_fishnet), 
    st_drop_geometry(pop_fishnet)) %>%
    st_join(studyAreaCounties) %>%
  filter(maj_water == FALSE)
dat <- dat %>%
  rename(lagDevelopment = Austin_fishnet.lagDevelopment) #attaching lagged development and renaming to fit in final dataset

tm_shape(dat) +
  tm_fill(col = "development_change2", palette = palette2, style = "cat")
```

## Exploratory Analysis
# In this section we try to understand to what extent a feature is associated with a cell being developed or not developed. We start with distance to highways and existing development. 
```{r, warning = FALSE, message = FALSE}
dat %>%
  dplyr::select(distance_highways,lagDevelopment,development_change2) %>%
  gather(Variable, Value, -development_change2, -geometry) %>%
  ggplot(., aes(development_change2, Value, fill=development_change2)) + 
    geom_bar(position = "dodge", stat = "summary", fun.y = "mean") +
    facet_wrap(~Variable) +
    scale_color_manual(values = palette2,
                      labels=c("No Change","New Development"),
                      name="") +
    labs(title="New Development as a Function of the Continuous Variables") +
    plotTheme 

# mutate to 1s and 0s
dat %>%
  dplyr::select(distance_highways, lagDevelopment, development_change2) %>%
  mutate(development_change2 = factor(development_change2, levels = c(0, 1))) %>%
  gather(Variable, Value, -development_change2, -geometry) %>%
  ggplot(., aes(development_change2, Value, fill = development_change2)) + 
  geom_bar(position = "dodge", stat = "summary", fun.y = "mean") +
  facet_wrap(~Variable) +
  scale_fill_manual(values = palette2,
                    labels = c("No Change", "New Development"),
                    name = "") +
  labs(title = "New Development as a Function of the Continuous Variables") +
  plotTheme

```

# Then we look at population. 
```{r, warning = FALSE, message = FALSE}
dat %>%
  dplyr::select(pop_2011,pop_2021,pop_Change,development_change2) %>%
  gather(Variable, Value, -development_change2, -geometry) %>%
  ggplot(., aes(development_change2, Value, fill=development_change2)) + 
    geom_bar(position = "dodge", stat = "summary", fun.y = "mean") +
    facet_wrap(~Variable) +
    scale_color_manual(values = palette2,
                      labels=c("No Change","New Development"),
                      name="") +
    labs(title="New Development as a Function of Factor Variables") +
    plotTheme

# mutate to 1s and 0s
dat %>%
  dplyr::select(pop_2011,pop_2021,pop_Change,development_change2) %>%
  mutate(development_change2 = factor(development_change2, levels = c(0, 1))) %>%
  gather(Variable, Value, -development_change2, -geometry) %>%
  ggplot(., aes(development_change2, Value, fill = development_change2)) + 
  geom_bar(position = "dodge", stat = "summary", fun.y = "mean") +
  facet_wrap(~Variable) +
  scale_fill_manual(values = palette2,
                    labels = c("No Change", "New Development"),
                    name = "") +
  labs(title = "New Development as a Function of the Continuous Variables") +
  plotTheme
```
# After visualizing our variables, we create a table of land cover conversion betweeen 2011 and 2021. 
```{r, warning = FALSE, message = FALSE}
dat %>%
  dplyr::select(development_change2:oUndev, dev, forest, water, wetland, farm) %>%
  gather(aggregatedRasters, Value, -development_change2, -geometry) %>%
   st_set_geometry(NULL) %>%
     group_by(development_change2, aggregatedRasters) %>%
     summarize(n = sum(as.numeric(Value))) %>%
     ungroup() %>%
    mutate(Conversion_Rate = paste0(round(100 * n/sum(n), 2), "%")) %>%
    filter(development_change2 == 1) %>%
  dplyr::select(aggregatedRasters, Conversion_Rate) %>%
  kable() %>% kable_styling(full_width = F)
```

# Predicting for 2021
# To do so, the data ('dat') is split into 50% training/test sets. 
```{r, warning = FALSE, message = FALSE}
set.seed(3456)
trainIndex <- 
  createDataPartition(dat$dev, p = .50,
                                  list = FALSE,
                                  times = 1)
datTrain <- dat[ trainIndex,]
datTest  <- dat[-trainIndex,]

nrow(dat)
```
# glm models estimated adding new variables 
```{r, warning = FALSE, message = FALSE}
Model1 <- glm(development_change2 ~ wetland + forest  + farm + oUndev, 
              family="binomial"(link="logit"), data = datTrain)

Model2 <- glm(development_change2 ~ wetland + forest  + farm + oUndev + lagDevelopment, 
              family="binomial"(link="logit"), data = datTrain)

Model3 <- glm(development_change2 ~ wetland + forest  + farm + oUndev + lagDevelopment + pop_2011,
              family="binomial"(link="logit"), data = datTrain)
              
Model4 <- glm(development_change2 ~ wetland + forest  + farm + oUndev + lagDevelopment + pop_2011 + pop_2021,
              family="binomial"(link="logit"), data = datTrain) 

Model5 <- glm(development_change2 ~ wetland + forest  + farm + oUndev + lagDevelopment + pop_Change,
              family="binomial"(link="logit"), data = datTrain) 
            
Model6 <- glm(development_change2 ~ wetland + forest  + farm + oUndev + lagDevelopment + pop_Change + distance_highways,
              family="binomial"(link="logit"), data = datTrain) 

```

```{r, warning = FALSE, message = FALSE}
modelList <- paste0("Model", 1:6)
map_dfc(modelList, function(x)pR2(get(x)))[4,] %>%
  setNames(paste0("Model",1:6)) %>%
  gather(Model,McFadden) %>%
  ggplot(aes(Model,McFadden)) +
    geom_bar(stat="identity") +
    labs(title= "McFadden R-Squared by Model") +
    plotTheme
```

# Model 4 shows the most promise.
```{r, warning = FALSE, message = FALSE}
testSetProbs <- 
  data.frame(class = datTest$development_change2,
             probs = predict(Model4, datTest, type="response")) 

mod_testSetProbs <- testSetProbs %>%
  dplyr::select(class,probs) %>%
  mutate(class = factor(class, levels = c(0, 1))) 
  
ggplot(mod_testSetProbs, aes(probs)) +
  geom_density(aes(fill=class), alpha=0.5) +
  scale_fill_manual(values = palette2,
                    labels=c("No Change","New Development")) +
  labs(title = "Histogram of test set predicted probabilities",
       x="Predicted Probabilities",y="Density") +
  plotTheme

```

# Accuracy
```{r, warning = FALSE, message = FALSE}
options(yardstick.event_first = FALSE)

testSetProbs <- 
  mod_testSetProbs %>% 
  mutate(predClass_05 = as.factor(ifelse(testSetProbs$probs >= 0.05 ,1,0)),
         predClass_17 = as.factor(ifelse(testSetProbs$probs >= 0.17 ,1,0))) 

testSetProbs %>% 
  dplyr::select(-probs) %>%
  gather(Variable, Value, -class) %>%
  group_by(Variable) %>%
  summarize(Sensitivity = round(yardstick::sens_vec(class,factor(Value)),2),
            Specificity = round(yardstick::spec_vec(class,factor(Value)),2),
            Accuracy = round(yardstick::accuracy_vec(class,factor(Value)),2)) %>% 
  kable() %>%
  kable_styling(full_width = F) 
```

```{r, warning = FALSE, message = FALSE}
predsForMap <-         
  dat %>%
    mutate(probs = predict(Model4, dat, type="response") ,
           Threshold_5_Pct = as.factor(ifelse(probs >= 0.05 ,1,0)),
           Threshold_17_Pct =  as.factor(ifelse(probs >= 0.17 ,1,0))) %>%
    dplyr::select(development_change2,Threshold_5_Pct,Threshold_17_Pct) %>%
    gather(Variable,Value, -geometry) %>%
    st_cast("POLYGON")
```

```{r, warning = FALSE, message= FALSE, fig.height = 6, fig.width= 8}
ggplot() +
  geom_point(data=predsForMap, aes(x=xyC(predsForMap)[,1], y=xyC(predsForMap)[,2], colour=Value)) +
  facet_wrap(~Variable) +
  scale_colour_manual(values = palette2, labels=c("No Change","New Development"),
                      name="") +
  labs(title="Development Predictions - Low Threshold") + 
  mapTheme
```

# Confusion Matrix
```{r, warning = FALSE, message = FALSE}
ConfusionMatrix.metrics <-
  dat %>%
    mutate(probs = predict(Model4, dat, type="response") ,
           Threshold_5_Pct = as.factor(ifelse(probs >= 0.05 ,1,0)),
           Threshold_17_Pct =  as.factor(ifelse(probs >= 0.17 ,1,0))) %>%
    mutate(TrueP_05 = ifelse(development_change2  == 1 & Threshold_5_Pct == 1, 1,0),
           TrueN_05 = ifelse(development_change2  == 0 & Threshold_5_Pct == 0, 1,0),
           TrueP_17 = ifelse(development_change2  == 1 & Threshold_17_Pct == 1, 1,0),
           TrueN_17 = ifelse(development_change2  == 0 & Threshold_17_Pct == 0, 1,0)) %>%
    dplyr::select(., starts_with("True")) %>%
    gather(Variable, Value, -geometry) %>%
    st_cast("POLYGON") 
```

# plot confusion matrix
```{r, warning = FALSE, message = FALSE, fig.height= 8, fig.width= 8 }
ggplot(data=ConfusionMatrix.metrics) +
  geom_point(aes(x=xyC(ConfusionMatrix.metrics)[,1], 
                 y=xyC(ConfusionMatrix.metrics)[,2], colour = as.factor(Value))) +
  facet_wrap(~Variable) +
  scale_colour_manual(values = palette2, labels=c("Correct","Incorrect"),
                       name="") +
  labs(title="Development Predictions - Low Threshold") + mapTheme
```

# generalizability 
```{r, warning = FALSE, message = FALSE}
spatialCV <- function(dataFrame, uniqueID, dependentVariable, modelName) {

#initialize a data frame 
endList <- list()

#create a list that is all the spatial group unqiue ids in the data frame (ie counties)    
  uniqueID_List <- unique(dataFrame[[uniqueID]])  
  x <- 1
  y <- length(uniqueID_List)
  
#create a counter and while it is less than the number of counties...  
  while(x <= y) 
  {
#call a current county    
    currentUniqueID <- uniqueID_List[x]
#create a training set comprised of units not in that county and a test set of units
#that are that county
    training <- dataFrame[ which(dataFrame[[uniqueID]] != uniqueID_List[x]),]
    testing <- dataFrame[ which(dataFrame[[uniqueID]] == uniqueID_List[x]),]
#create seperate xy vectors
    trainingX <- training[ , -which(names(training) %in% c(dependentVariable))]
    testingX <- testing[ , -which(names(testing) %in% c(dependentVariable))]
    
    trainY <- training[[dependentVariable]]
    testY <- testing[[dependentVariable]]
#Calculate predictions on the test county as part of a data frame including the observed
#outcome and the unique county ID    
   thisPrediction <- 
     data.frame(class = testY,
                probs = predict(modelName, testingX, type="response"),
                county = GEOID) 

#Row bind the predictions to a data farme
   endList <- rbind(endList, thisPrediction)
#iterate counter    
    x <- x + 1 
  } 
#return the final list of counties and associated predictions  
  return (as.data.frame(endList))
}
```

```{r, warning = FALSE, message = FALSE}

spatialCV_counties <-
  spatialCV(dat3,"NAME","development_change2", Model4) %>%
  mutate(predClass = as.factor(ifelse(probs >= 0.17 ,1,0))) 

dat3 <- unique(dat3)

testSetProbs <- 
  data.frame(class = datTest$development_change2,
             probs = predict(Model4, datTest, type="response")) 

dat3 <- dat3 %>%
  dplyr::select(class,probs) %>%
  mutate(class = factor(class, levels = c(0, 1))) 
  
ggplot(mod_testSetProbs, aes(probs)) +
  geom_density(aes(fill=class), alpha=0.5) +
  scale_fill_manual(values = palette2,
                    labels=c("No Change","New Development")) +
  labs(title = "Histogram of test set predicted probabilities",
       x="Predicted Probabilities",y="Density") +
  plotTheme


# adding county boundaries to fishnet
add_pop_county21 <- get_acs(
  geography = "county", variables = "B01001_001", state = 48,  geometry = TRUE, 
                county=c("Hays", "Travis", "Williamson","Bastrop","Caldwell"), year = 2021) %>% 
  st_transform(crs = 6578) %>%
  rename(pop_2021 = estimate) %>% 
  st_buffer(-1)

ggplot() +
  geom_sf(data = pop_county21)
view(pop_county21)

# adding county boundaries to dat3 (made a new version of dat)

dat3 <- st_join(dat, pop_county21, by = "pop_21", all = TRUE)
ggplot(data = dat3) +
  geom_sf(aes(fill = GEOID)) +
  scale_fill_manual(values = palette5) 

```
```{r, warning = FALSE, message = FALSE}
spatialCV_metrics <-
  spatialCV_counties %>% 
    group_by(county) %>% 
    summarize(Observed_Change = sum(as.numeric(as.character(class))),
              Sensitivity = round(yardstick::sens_vec(class,predClass),2),
              Specificity = round(yardstick::spec_vec(class,predClass),2),
              Accuracy = round(yardstick::accuracy_vec(class,predClass),2)) 

spatialCV_metrics %>%
  kable() %>%
  kable_styling(full_width = F)
```

## Scenario 1: Predicting for 2031
```{r, warning = FALSE, message = FALSE}
dat <-
  dat %>%
  mutate(lagDevelopment = nn_function(xyC(.), xyC(filter(.,developed2021 == 1)),2))
```

# add in population data (from: https://demographics.texas.gov/Projections/2022/)
```{r, warning = FALSE, message = FALSE}
# editing pop_2021 table
pop_2021$county <- str_extract(pop_2021$NAME, "[A-Z][a-z]+ County")
  
countyPopulation_2031 <- 
  data.frame(
   county = 
     c("Hays County", "Travis County", "Williamson County","Bastrop County","Caldwell County"),
   county_projection_2031 = 
     c(369690,1599770,838036,123564,54523)) 

pop_2031 <- countyPopulation_2031 %>%
   left_join(
     pop_2021 %>%
       st_set_geometry(NULL) %>%
       group_by(county) %>%
       summarize(county_pop_2021 = sum(pop_2021)))

pop_2031 %>%
  gather(Variable,Value, -county) %>%
  ggplot(aes(reorder(county,-Value),Value)) +
  geom_bar(aes(fill=Variable), stat = "identity", position = "dodge") +
  scale_fill_manual(values = palette2,
                    labels=c("2031","2021"),
                    name="Population") +
  labs(title="Population Change by County: 2021 - 2031",
       x="County", y="Population") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  plotTheme
```

# Predicting Development Demand
```{r, warning = FALSE, message = FALSE}
dat_infill <-
  dat %>%
  #calculate population change
    cross_join(pop_2031) %>%
    mutate(proportion_of_county_pop = pop_2021/ sum(pop_2021),
           pop_2031.infill = proportion_of_county_pop * county_projection_2031,
           pop_Change = round(pop_2031.infill - county_pop_2021),2) %>%
    dplyr::select(-county_projection_2031, -county_pop_2021, 
                  -proportion_of_county_pop, -pop_2031.infill) %>%
  #predict for 2031
    mutate(predict_2031.infill = predict(Model4,. , type="response"))

# to visualize by county 
MSACounty <- get_acs(
  geography = "county", variables = "B01001_001", state = 48,  geometry = TRUE, 
                county=c("Hays", "Travis", "Williamson","Bastrop","Caldwell"), year = 2021) %>% 
  st_transform(crs = 6578) %>%
  rename(pop_2021 = estimate) %>% 
  st_buffer(-1)

dat_infill %>%
  ggplot() +  
  geom_point(aes(x=xyC(dat_infill)[,1], y=xyC(dat_infill)[,2], colour = factor(ntile(predict_2031.infill,5)))) +
  scale_colour_manual(values = palette5,
                    labels=substr(quintileBreaks(dat_infill,"predict_2031.infill"),1,4),
                    name="Quintile\nBreaks") +
   geom_sf(data=MSACounty, fill=NA, colour="black", size=5) +
  labs(title= "Development Demand in 2031: Predicted Probabilities") +
  mapTheme 

```

# Sensitive Land Cover Lost
```{r, warning = FALSE, message = FALSE, fig.height = 6, fig.width= 6}

dat2 <- dat
dat2 <- dat2 %>% 
  mutate(sensitive_lost = ifelse(development_change2 == 1 & forest >= 0.5 |
                                   development_change2 == 1 & wetland >= 0.5,1,0))
ggplot() +
  geom_point(data=dat2, aes(x=xyC(dat2)[,1], y=xyC(dat2)[,2], 
                            colour=as.factor(sensitive_lost))) +
  scale_colour_manual(values = palette2,
                      labels=c("No Change","Sensitive Lost"),
                      name = "") +
  labs(title = "Sensitive lands lost: 2011 - 2021",
       subtitle = "As fishnet centroids") +
  mapTheme

# Creating sensitive regions
sensitiveRegions <- 
  raster::clump(wetland + forest) %>%
  rasterToPolygons() %>%
  st_as_sf() %>%
  group_by(clumps) %>% 
  summarize() %>%
    mutate(Acres = as.numeric(st_area(.) * 0.0000229568)) %>%
    filter(Acres > 3954)  %>%
  dplyr::select() %>%
  raster::rasterize(.,emptyRaster) 
sensitiveRegions[sensitiveRegions > 0] <- 1  
names(sensitiveRegions) <- "sensitiveRegions"

dat2 <-
  aggregateRaster(c(sensitiveRegions), dat2) %>%
  dplyr::select(sensitiveRegions) %>%
  st_set_geometry(NULL) %>%
  bind_cols(.,dat2) %>%
  st_sf() 

ggplot() +
  geom_point(data=dat2, aes(x=xyC(dat2)[,1], y=xyC(dat2)[,2], colour=as.factor(sensitiveRegions))) +
  scale_color_manual(values = palette2,
                      labels=c("Other","Sensitive Regions"),
                      name="") +
  labs(title = "Sensitive regions",
       subtitle = "Continous areas of either wetlands or forests\ngreater than 1 acre") +
  mapTheme # NOT ENTIRELY SURE ABOUT THIS MAP, feels like too many sensitive regions?

# summarize by county 
county_specific_metrics <- 
  dat2 %>%
  #predict development demand from our model
  mutate(Development_Demand = predict(Model4, dat2, type="response")) 
  #get a count count of grid cells by county which we can use to calculate rates below




```

# Allocation # CURRENTLY WORKING
```{r, warning = FALSE, message = FALSE, fig.height= 8, fig.width= 11}
travis_county <-
  county_specific_metrics %>%
    mutate(Development_Demand = predict(Model4, dat2, type="response")) %>%
    filter(NAME == "Travis County") 

fortBend_landUse <- rbind(
  filter(fortBend, forest11 == 1 | wetlands11 == 1 ) %>%
  dplyr::select() %>% mutate(Land_Use = "Not Suitable"),
  filter(fortBend, developed11 == 1) %>%
  dplyr::select() %>% mutate(Land_Use = "Developed"))

grid.arrange(
ggplot() +
  geom_sf(data=fortBend, aes(fill=factor(ntile(Development_Demand,5))), colour=NA) +
  geom_point(data=fortBend_landUse, aes(x=xyC(fortBend_landUse)[,1], 
                                        y=xyC(fortBend_landUse)[,2], colour=Land_Use),
                                        shape = 15, size = 2) +
  geom_sf(data=st_intersection(houstonHighways,filter(studyAreaCounties, NAME=="Fort Bend")), size=2) +
  scale_fill_manual(values = palette5, name="Development\nDemand",
                    labels=substr(quintileBreaks(fortBend,"Development_Demand"),1,5)) +
  scale_colour_manual(values = c("black","red")) + 
  labs(title = "Development Potential, 2020: Fort Bend") + mapTheme +
  guides(fill = guide_legend(order = 1), colour = guide_legend(order = 2)),

ggplot() +
  geom_sf(data=fortBend, aes(fill=factor(ntile(pop_Change,5))), colour=NA) +
  geom_point(data=fortBend_landUse, aes(x=xyC(fortBend_landUse)[,1], 
                                        y=xyC(fortBend_landUse)[,2], colour=Land_Use),
                                        shape = 15, size = 2) +
  geom_sf(data=st_intersection(houstonHighways,filter(studyAreaCounties, NAME=="Fort Bend")), size=2) +
  scale_fill_manual(values = palette5, name="Population\nChange",
                    labels=substr(quintileBreaks(fortBend,"pop_Change"),1,5)) +
  scale_colour_manual(values = c("black","red")) + 
  labs(title = "Projected Population, 2020: Fort Bend") + mapTheme +
  guides(fill = guide_legend(order = 1), colour = guide_legend(order = 2)), ncol=2)
```


## Scenario 2: Estimating the effect of new transportation
```{r, warning = FALSE, message = FALSE}
library(leaflet)
library(leafem)

# Create a leaflet map
newhwy <- leaflet() %>%
  addTiles(MSAHighways) 
 
# Add highway infrastructure data as map layers
# Replace `highway_data` with your own highway infrastructure data
newhwy_map <- newhwy %>%
  addPolygons(data = MSAHighways) %>%
  addPolylines(data = MSAHighways)

# Add editing controls
newhwy_map <- newhwy_map %>%
  leaflet::addDrawToolbar(targetGroup='edit')

# Display the map
newhwy_map

```
